\documentclass[11pt, twoside, openright, utf8, bachelor, english, subfiles]{base/thesis}

\usepackage[english]{babel}
%\usepackage{cite}
\usepackage{natbib}
\bibpunct{(}{)}{;}{a}{,}{,}
\bibliographystyle{agsm}

\usepackage{tikz}    

\usepackage{listings}
\renewcommand{\lstlistingname}{Listing}

\usepackage[]{algorithm2e}

\usepackage{graphicx}
%\usepackage{subcaption}
%\usepackage[export]{adjustbox}
%\usepackage{sidecap} % for captions on the side of figures

\usepackage{wrapfig}

\usepackage{ragged2e}

% This might or might not be useful. It ensures, that latex doesn't overly stretch whitespaces between paragraphs, because it thinks that would be nice.
%\raggedbottom


% for the pdf meta data to work properly, these have to be set in the preamble
\institute{Institute of Cognitive Science}
% \workgroup{<workgroup>}
\title{A textual recommender system and other text mining applications for clinical data}
\author{Philipp Hummel}
\city{Osnabrück}
\submitdate{\today}
\firstSupervisor{Dr. Frank Jäkel}
\secondSupervisor{Prof. Dr. Gordon Pipa}

\begin{document}

\maketitle

\begin{abstract}
\subfile{chapters/abstract}
\end{abstract}

\begin{acknowledgements}
I would like to thank several people and acknowledge their contributions to this thesis. First and foremost I want to thank Dr. Frank Jäkel for his guidance, his expertise in cognitive science, psychology and machine learning, and generally his well thought-out remarks. Thanks also to Dr. Sascha Lange and the company PSIORI for making this thesis possible by selecting me for this project, providing the objective for the thesis, and sharing their contacts and resources.

Special thanks go to Prof. Dr. Roland Mertelsmann who envisioned this project for years, gave many fruitful comments, established the connection to the University hospital Freiburg and motivated several other physicians and students to contribute to this project among other things by participating in the experiment.

Lastly I want to thank Astrid Feiten, Judith Schepers and Lisa Dieminger for their proof-reading, that went far beyond plain grammar correction.
\end{acknowledgements}

\tableofcontents

\mainpart
\subfile{chapters/introduction}
\subfile{chapters/dataset_cleansing}
\subfile{chapters/fine-grained_information_presentation}
\subfile{chapters/recommender_system}
\subfile{chapters/intelligent_search}
\subfile{chapters/discussion}
%\subfile{chapters/results}
%\subfile{chapters/experiment}
%\subfile{chapters/prototype}
%\subfile{chapters/conclusion}

\appendix
\subfile{chapters/appendix}
% appendixcontents: 
%\listoffigures

\bibliographystyle{apalike}
\bibliography{thesis_bibliography}


\declaration

\end{document}